Investigating motor vehicle collisions in Toronto: Installing more red light cameras do not stop traffic accidents from happening

### 2022-02-06 outputs/paper/.ipynb_checkpoints/paper-checkpoint.Rmd
discussion and introduction

The Motor Vehicle Collisions data of Toronto provides a record of the past severe vehicle collisions in Toronto which result in injuries or deaths, and with the help of statistical data analysis, it can possibly offer insights on the ways to reduce this type of accident. This report investigates the motor vehicle collisions in Toronto from 2006 to 2020, with a focus on the ones result in injuries or deaths. In Section \@ref(data), the Motor Vehicle Collisions dataset is introduced and organized to visualize the change in the annual number of traffic accidents with respect to their fatality and their location. Then, I discussed about the possible implications of these visualizations. By Figure 2, I concluded that motor vehicle collisions are always more likely to occur at intersections. Thus, 
Continuing

You can and should cross-reference sections and sub-sections. For instance, Section \@ref(data). R Markdown automatically makes the sections lower case and adds a dash to spaces to generate labels, for instance, Section \@ref(first-discussion-point).

# Data
## Motor Vehicle Collisions involving Killed or Seriously Injured Persons
To investigate the motor vehicle collisions in Toronto, I obtained the dataset "Motor Vehicle Collisions involving Killed or Seriously Injured Persons" from the Toronto Open Data Portal [@citedata]. Provided by the Toronto Police Service, this dataset includes all vehicle collisions occurred in Toronto between 2006 and 2020 that result in injuries or deaths. One bias aroused by this dataset is the offset from the real accident location. For ethical reason and the protection of the parties involved in the accidents that are considered to be criminal cases, the location of those criminal accidents are offset to the closest intersection node [@citecollisiondata]. As a result, the location of those accidents specific to neighborhoods might not be accurate. Therefore, I will use the variable district instead of neighbourhood to categorize the accidents later in the report to avoid this bias.

@@

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Load motor vehicle collision data and clean
vehicle_collision <- read_csv(here::here("inputs/data/vehicle_collisions.csv")) %>%
  clean_names() %>%
  drop_na(year) %>%
  filter(acclass != "Property Damage Only")
```

@@

```{r fig.cap = "Comparing the number of traffic collisions each year from 2006 to 2020", fig.width=8.4,fig.height=3.6, echo=FALSE}
# Create a bar plot to show number of collisions by fatality
ggplot(mapping = aes(x = year, fill = acclass), data=vehicle_collision) +
  geom_bar()+theme_light()+ylab("number of motor vehicle collisions")+scale_fill_discrete(name = "Fatality")+facet_wrap(~acclass)
```

@@

```{r fig.cap = "Comparing the location of traffic collisions each year from 2006 to 2020", echo=FALSE, fig.width=6,fig.height=3.2}
# Create a bar plot to show number of collisions by accident location
ggplot(mapping = aes(x = year, fill=accloc), data=vehicle_collision)+scale_fill_discrete(name = "Accident location")+ylab("number of motor vehicle collisions")+theme_light()+geom_bar()
```

@@

```{r dadaextract1, message=FALSE, warning=FALSE, echo=FALSE}
# Load red light camera data and clean
red_light_camera <- read_csv(here::here("inputs/data/red_light_camera.csv")) %>%
  clean_names() %>%
  select(activation_date, district) %>%
  drop_na(activation_date) %>%
  drop_na(district)

@@

```{r message=FALSE, warning=FALSE, echo=FALSE, fig.cap = "Number of red light cameras in the four districts each year", fig.width=7,fig.height=3.2}
# Create a new dataset that presents the total number of red light cameras in each district
red_light_camera <- red_light_camera %>%
  filter(year != 2021)
cumulative <-tibble(
  year=c(2006, 2007, 2008, 2009, 2010, 2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018, 2019, 2020, 2006, 2007, 2008, 2009, 2010, 2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018, 2019, 2020, 2006, 2007, 2008, 2009, 2010, 2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018, 2019, 2020, 2006, 2007, 2008, 2009, 2010, 2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018, 2019, 2020),
  sum=c(0, 2, 6, 15, 16, 17, 17, 17, 17, 17, 17, 31, 33, 33, 34, 0, 0, 7, 17, 18, 18, 18, 18, 18, 18, 18, 42, 43, 44, 44, 0, 3, 10, 18, 19, 19, 19, 19, 19, 19, 19, 26, 27, 28, 28, 0, 1, 9, 15, 15, 15, 15, 15, 15, 15, 15, 36, 38, 41, 41),
  district=c("Toronto and East York", "Toronto and East York", "Toronto and East York", "Toronto and East York", "Toronto and East York", "Toronto and East York", "Toronto and East York", "Toronto and East York", "Toronto and East York", "Toronto and East York", "Toronto and East York", "Toronto and East York", "Toronto and East York", "Toronto and East York", "Toronto and East York", "Scarborough", "Scarborough", "Scarborough", "Scarborough", "Scarborough", "Scarborough", "Scarborough", "Scarborough", "Scarborough", "Scarborough", "Scarborough", "Scarborough", "Scarborough", "Scarborough", "Scarborough", "North York", "North York", "North York", "North York", "North York", "North York", "North York", "North York", "North York", "North York", "North York", "North York", "North York", "North York", "North York", "Etobicoke York", "Etobicoke York", "Etobicoke York", "Etobicoke York", "Etobicoke York", "Etobicoke York", "Etobicoke York", "Etobicoke York", "Etobicoke York", "Etobicoke York", "Etobicoke York", "Etobicoke York", "Etobicoke York", "Etobicoke York", "Etobicoke York")
)

@@

I wanted to see how the increase in number of red light cameras compare with the traffic collision's frequency in each district. So, I selected the accidents that occurred at intersections or somewhere near intersections from the Motor Vehicle Collision dataset and created a ggplot to see their trend by each district in Toronto.

```{r, echo=FALSE, warning=FALSE, fig.cap = "Comparing the number of motor vehicle collisions take place at or near intersections by district", fig.width=7,fig.height=3.2}
# Select the accidents that only occur at or near intersections
co <- vehicle_collision %>%
  filter(accloc == "At Intersection" | accloc == "Intersection Related") %>%
  drop_na(district)
for(i in 1:9039){
  if(co$district[i] == "Toronto East York"){
    co$district[i] = "Toronto and East York"
  }
}

@@

# Discussion
The analysis of the two datasets show no relationship between the cases of motor vehicle collisions in Toronto and the number of red light cameras. From the comparision between Figure 3 and Figure 4, we see that in none of the districts did the number of vehicle collisions followed the trend of the installations of the red light cameras. For stance, between 2016 and 2017, all four districts made the installations of lots of new red light cameras, but none of the data in Figure 4 seemed to reflect that action. The implication of Section \@ref(data) is that: installing more red cameras does not change the number of traffic accidents in Toronto, nor does it influence the fatality of the accidents. A very likely reason to explain this situation is that red light cameras only concerned the drivers who already followed the rule most of the time, but not the rule-breaking drivers. Another possible reason would be that at intersections with red light cameras, more drivers would intend to stop rather than to proceed when they see the light is about to turn red, resulting in a higher chance of rear end accidents [@cite2].

>>>>

The Motor Vehicle Collisions data of Toronto provides a record of the past severe vehicle collisions in Toronto which result in injuries or deaths, and with the help of statistical data analysis, it can possibly offer insights on the ways to reduce this type of accident. This report investigates the motor vehicle collisions in Toronto from 2006 to 2020, with a focus on the ones result in injuries or deaths. In Section \@ref(motor-vehicle-collisions-involving-killed-or-seriously-injured-persons), the Motor Vehicle Collisions dataset is introduced and organized to visualize the change in the annual number of traffic accidents with respect to their fatality and their location. Then, I discussed about the possible implications of these visualizations. Figure 1 showed that there was a descending trend of traffic accidents from 2006 to 2020. By Figure 2, I concluded that motor vehicle collisions are always more likely to occur at intersections. Thus, I brought out the Red Light Cameras dataset in Section \@ref(red-light-cameras) to see whether if the number of red light cameras has an effect on the number of traffic accidents with serious injuries and deaths. The result from the datasets indicates that there is no clear relationship between this two factors, suggesting that just by installing more red light cameras cannot prevent traffic accidents from happening at intersections. More implications and further discussions on this result is presented in Section \@ref(discussion).

# Data
## Motor Vehicle Collisions involving Killed or Seriously Injured Persons
To investigate the motor vehicle collisions in Toronto, I obtained the dataset "Motor Vehicle Collisions involving Killed or Seriously Injured Persons" from the Toronto Open Data Portal [@citedata]. Provided by the Toronto Police Service, this dataset includes all vehicle collisions occurred in Toronto between 2006 and 2020 that result in injuries or deaths. One bias aroused by this dataset is the offset from the real accident location. For ethical reason and the protection of the parties involved in the accidents that are considered to be criminal cases, the location of those criminal accidents are offset to the closest intersection node [@citecollisiondata]. As a result, the location of those accidents specific to neighborhoods might not be accurate. Therefore, I will use the variable district instead of the variable neighbourhood to categorize the accidents later in the report to avoid the bias.

@@

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Load motor vehicle collision data and clean
vehicle_collision <- read_csv(here::here("inputs/data/vehicle_collisions.csv")) %>%
  clean_names() %>%
  drop_na(year) %>%
  filter(acclass != "Property Damage Only")
```

@@

```{r echo=FALSE, fig.cap="Comparing the number of traffic collisions each year from 2006 to 2020", fig.width=8.4, fig.height=3.6}
# Create a bar plot to show number of collisions by fatality
ggplot(mapping = aes(x = year, fill = acclass), data=vehicle_collision) +
  geom_bar()+theme_light()+ylab("number of motor vehicle collisions")+scale_fill_discrete(name = "Fatality")+facet_wrap(~acclass)
```

@@

```{r echo=FALSE, fig.cap="Comparing the location of traffic collisions each year from 2006 to 2020", fig.width=6, fig.height=3.2}
# Create a bar plot to show number of collisions by accident location
ggplot(mapping = aes(x = year, fill=accloc), data=vehicle_collision)+scale_fill_discrete(name = "Accident location")+ylab("number of motor vehicle collisions")+theme_light()+geom_bar()
```

@@

```{r dadaextract1, echo=FALSE, message=FALSE, warning=FALSE}
# Load red light camera data and clean
red_light_camera <- read_csv(here::here("inputs/data/red_light_camera.csv")) %>%
  clean_names() %>%
  select(activation_date, district) %>%
  drop_na(activation_date) %>%
  drop_na(district)

@@

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Number of red light cameras in the four districts each year", fig.width=7, fig.height=3.2}
# Create a new dataset that presents the total number of red light cameras in each district
red_light_camera <- red_light_camera %>%
  filter(year != 2021)
cumulative <-tibble(
  year=c(2006, 2007, 2008, 2009, 2010, 2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018, 2019, 2020, 2006, 2007, 2008, 2009, 2010, 2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018, 2019, 2020, 2006, 2007, 2008, 2009, 2010, 2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018, 2019, 2020, 2006, 2007, 2008, 2009, 2010, 2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018, 2019, 2020),
  sum=c(0, 2, 6, 15, 16, 17, 17, 17, 17, 17, 17, 31, 33, 33, 34, 0, 0, 7, 17, 18, 18, 18, 18, 18, 18, 18, 42, 43, 44, 44, 0, 3, 10, 18, 19, 19, 19, 19, 19, 19, 19, 26, 27, 28, 28, 0, 1, 9, 15, 15, 15, 15, 15, 15, 15, 15, 36, 38, 41, 41),
  district=c("Toronto and East York", "Toronto and East York", "Toronto and East York", "Toronto and East York", "Toronto and East York", "Toronto and East York", "Toronto and East York", "Toronto and East York", "Toronto and East York", "Toronto and East York", "Toronto and East York", "Toronto and East York", "Toronto and East York", "Toronto and East York", "Toronto and East York", "Scarborough", "Scarborough", "Scarborough", "Scarborough", "Scarborough", "Scarborough", "Scarborough", "Scarborough", "Scarborough", "Scarborough", "Scarborough", "Scarborough", "Scarborough", "Scarborough", "Scarborough", "North York", "North York", "North York", "North York", "North York", "North York", "North York", "North York", "North York", "North York", "North York", "North York", "North York", "North York", "North York", "Etobicoke York", "Etobicoke York", "Etobicoke York", "Etobicoke York", "Etobicoke York", "Etobicoke York", "Etobicoke York", "Etobicoke York", "Etobicoke York", "Etobicoke York", "Etobicoke York", "Etobicoke York", "Etobicoke York", "Etobicoke York", "Etobicoke York")
)

@@

I wanted to see how the increase in number of red light cameras compare with the traffic collision's frequency in each district. So, I selected the accidents that occurred at intersections or somewhere near intersections from the Motor Vehicle Collision dataset and created a bar plot to see their trend by each district in Toronto.

```{r echo=FALSE, warning=FALSE, fig.cap="Comparing the number of motor vehicle collisions take place at or near intersections by district", fig.width=7, fig.height=3.2}
# Select the accidents that only occur at or near intersections
co <- vehicle_collision %>%
  filter(accloc == "At Intersection" | accloc == "Intersection Related") %>%
  drop_na(district)
for(i in 1:9039){
  if(co$district[i] == "Toronto East York"){
    co$district[i] = "Toronto and East York"
  }
}

@@

# Discussion
The analysis of the two datasets show no relationship between the cases of motor vehicle collisions in Toronto and the number of red light cameras. From the comparison between Figure 3 and Figure 4, we see that in none of the districts did the number of vehicle collisions followed the trend of the installations of the red light cameras. For stance, between 2016 and 2017, all four districts made the installations of lots of new red light cameras, but none of the data in Figure 4 seemed to reflect that action. The implication of Section \@ref(data) is that: installing more red cameras does not change the number of traffic accidents in Toronto, nor does it influence the fatality of the accidents. A very likely reason to explain this situation is that red light cameras only concerned the drivers who already followed the rule most of the time, but not the rule-breaking drivers. Another possible reason would be that at intersections with red light cameras, more drivers would intend to stop rather than to proceed when they see the light is about to turn red, resulting in a higher chance of rear end accidents [@cite2].
